---
title: "S. angularis Landscape Genetic analyses"
author: "Sarah Emel"
date: "6/6/2020"
output: pdf_document
---


Load libraries
```{r}
library(MuMIn)
library(lme4)
library(rdist)
library(psych)
```

## MLPE functions
```{r}
#MLPE functions
#use MLPE for model fitting, examining coefficients, CIs
#use MLPEnoREML for model comparison

MLPE <- function(variables, data) {
  mod2 <- lme4::lFormula(variables, data = data, REML = TRUE)
  dfun <- do.call(lme4::mkLmerDevfun, mod2)
  opt <- lme4::optimizeLmer(dfun)
  mod_2 <- lme4::mkMerMod(environment(dfun), opt, mod2$reTrms,fr = mod2$fr)
  mod2$reTrms$Zt <- ZZ
  # Refit the model
  dfun <- do.call(lme4::mkLmerDevfun, mod2)
  opt <- lme4::optimizeLmer(dfun)
  modelout <- lme4::mkMerMod(environment(dfun), opt, mod2$reTrms,fr = mod2$fr)
  return(modelout)
}

MLPEnoREML <- function(variables, data) {
  mod2 <- lme4::lFormula(variables, data = data, REML = FALSE)
  dfun <- do.call(lme4::mkLmerDevfun, mod2)
  opt <- lme4::optimizeLmer(dfun)
  mod_2 <- lme4::mkMerMod(environment(dfun), opt, mod2$reTrms,fr = mod2$fr)
  mod2$reTrms$Zt <- ZZ
  # Refit the model
  dfun <- do.call(lme4::mkLmerDevfun, mod2)
  opt <- lme4::optimizeLmer(dfun)
  modelout <- lme4::mkMerMod(environment(dfun), opt, mod2$reTrms,fr = mod2$fr)
  return(modelout)
}
```

## Import LCP and Circuitscape output

*Double check filepaths!*
```{r}
library(foreign)

#get list of lcp files
lcp.list <- list.files("./Sabatia_Resistance_2Rules_out/", pattern = ".dbf", full.names = T)
test <- read.dbf(lcp.list[[1]]) #get first file
df_lcp <- data.frame(test$pair_name) #use first file to create dataframe

#add columns to dataframe for each file in list
for (i in 1:length(lcp.list)){
  x <- read.dbf(lcp.list[[i]])
  y <- x[,2]
  df <- data.frame(y) 
}

#format names
lcpnames <- gsub("./Sabatia_Resistance_2Rules_out//new", lcp.list, fixed = TRUE, replacement = "LCP")
lcpnames <- gsub(".dbf", "", lcpnames, fixed = TRUE)
colnames(df_lcp) <- c("pair",lcpnames)
#write.csv(df_lcp, "Sang_LCP_2Rules.csv") #uncomment to write file

#get list of circuitscape files
cs.list <- list.files("./forCircuitscape_2Rules/", pattern = "_3columns", full.names = T)
test <- read.table(cs.list[[1]]) #get first file
df_cs <- data.frame(test[,1:2]) #use first file to create dataframe

#add columns to dataframe for each file in list
for (i in 1:length(cs.list)){
  x <- read.table(cs.list[[i]])
  y <- x[,3]
  df <- data.frame(y)
  df_cs <- cbind(df_cs,df)
}

#format names
csnames <- gsub("./forCircuitscape_2Rules//newcost", cs.list, fixed = TRUE,replacement = "resist")
csnames <- gsub("_resistances_3columns.asc", "", csnames, fixed = TRUE)
csnames <- gsub("_resistances_3columns.out", "", csnames, fixed = TRUE)
colnames(df_cs) <- c("pop1","pop2",csnames)
#write.csv(df_cs, "Sang_CS_2Rules.csv") #uncomment to write file

#note: writing files for convenience
#dataframe(s) could be used directly in next step
```


## Read in all files needed

Read in data: genetic distance, IBD, IBR, IBE
```{r}
#genetic distance: FST
fst <- read.csv("Sang_Fst_Genodive.csv", header = T)
gdist <- fst[lower.tri(fst)] #get lower triangle of matrix

#populations: make sure all pops listed in both pop1 and pop2
pop <- read.csv(file="pops.csv",header=T)
pop <- pop[,c(2:5,7)]

#IBD, IBR:LCP and Circuitscape
lcp <- read.csv(file="Sang_LCP_2Rules.csv",header=T)
lcp <- lcp[,-c(1:2)]
cs <- read.csv(file="Sang_CS_2Rules.csv",header=T)
cs <- cs[,-c(1:3)]

#soil data
soil <- read.csv(file="envfinalall.csv", header=T) #site means

#check that things are in the same order
as.vector(soil$Pop) == colnames(fst)
```

## Calculate IBE

These are done for each metal separately, but could be combined with PCA
```{r}
#Cr
Cr <- as.data.frame(soil[,c("Cr")])
row.names(Cr)<-soil$Pop #changing population names column to row names
IBE.Cr<-data.matrix(rdist(Cr))
IBE.Cr<-IBE.Cr[lower.tri(IBE.Cr)]
#Cu
Cu <- as.data.frame(soil[,c("Cu")])
row.names(Cu)<-soil$Pop #changing population names column to row names
IBE.Cu<-data.matrix(rdist(Cu))
IBE.Cu<-IBE.Cu[lower.tri(IBE.Cu)]
#Ni
Ni <- as.data.frame(soil[,c("Ni")])
row.names(Ni)<-soil$Pop #changing population names column to row names
IBE.Ni<-data.matrix(rdist(Ni))
IBE.Ni<-IBE.Ni[lower.tri(IBE.Ni)]
#Zn
Zn <- as.data.frame(soil[,c("MehZn")])
row.names(Zn)<-soil$Pop #changing population names column to row names
IBE.Zn<-data.matrix(rdist(Zn))
IBE.Zn<-IBE.Zn[lower.tri(IBE.Zn)]
```

## Combine and prep data
Combine IBD, IBR, and IBE data into a single dataframe
```{r}
alldata <- cbind(gdist,pop,lcp,cs,IBE.Cr,IBE.Cu,IBE.Ni,IBE.Zn)
```

## Run MLPE for all pairs of sites

Scale and center distances
```{r}
all.scaled<-scale(alldata[,6:ncol(alldata)],center=T,scale=T)
all.scaled<-cbind(alldata[,1:5],all.scaled)

data<-all.scaled
head(alldata)
pop1<-data$pop1
pop2<-data$pop2
```

Create output file and run MLPE functions
```{r}
#heading for the output file
header <- noquote(paste("Var","AICc", "BIC", "R2m", "R2c","Est","2.5%", "97.5%",sep=" ")) 
write.table(header, file = "Sang_MLPE_results_2Rules_alldata.txt", append=T, row.names=F, col.names=F, quote = F) 

#prep
Zl <- lapply(c("pop1","pop2"), function(nm) Matrix:::fac2sparse(data[[nm]],"d", drop=FALSE))
ZZ <- Reduce("+", Zl[-1], Zl[[1]])

#run MLPE loop (always check starting column #)
for (i in (6:ncol(data))) {
  #do model comparison with MLPEnoREML
  fit <- MLPEnoREML(gdist ~ data[, i] + (1|pop1), data=data)
    anova(fit)
    a <- AICc(fit, k = length(unique(pop1))) #this make k = # pops, not # pairs
    b <- BIC(fit)
    r <- r.squaredGLMM(fit)
    rm <- r[1]
    rc <- r[2]
  #get parameter estimates with REML
  fitREML <- MLPE(gdist ~ data[, i] + (1|pop1), data=data)
    est <- coef(summary(fitREML))
    ci <- confint(fitREML, level = 0.95, method = "Wald")
  out<- noquote(paste(colnames(data)[i],a, b, rm, rc,est[2,1], ci[4,1], ci[4,2],
                      sep =" "))
  #write.table(out, file = "Sang_MLPE_results_2Rules_alldata.txt", append=T,row.names=F,col.names=F, quote = F) #uncomment to write file
}
```


## Run MLPE for within-spatial cluster only

Subset for within-spatial cluster only and scale distances
```{r}
sepdata<-alldata[which(alldata$Cluster=='S'|alldata$Cluster=='N'),]
data<-sepdata

sep.scaled<-scale(sepdata[,6:ncol(sepdata)],center=T,scale=T)
sep.scaled<-cbind(sepdata[,1:5],sep.scaled)

data<-sep.scaled
pop1<-data$pop1
pop2<-data$pop2
```

Create output file and run MLPE functions
```{r}
#heading for the output file
header <- noquote(paste("Var","AICc", "BIC", "R2m", "R2c","Est","2.5%", "97.5%",sep=" ")) 
write.table(header, file = "Sang_MLPE_results_2Rules_separate.txt", append=T, row.names=F, col.names=F, quote = F) 

#prep
Zl <- lapply(c("pop1","pop2"), function(nm) Matrix:::fac2sparse(data[[nm]],"d", drop=FALSE))
ZZ <- Reduce("+", Zl[-1], Zl[[1]])

#run MLPE loop (always check starting column #)
for (i in (6:ncol(data))) {
  #do model comparison with MLPEnoREML
  fit <- MLPEnoREML(gdist ~ data[, i] + (1|pop1), data=data)
    anova(fit)
    a <- AICc(fit, k = length(unique(pop1))) #this make k = # pops, not # pairs
    b <- BIC(fit)
    r <- r.squaredGLMM(fit)
    rm <- r[1]
    rc <- r[2]
  #get parameter estimates with REML
  fitREML <- MLPE(gdist ~ data[, i] + (1|pop1), data=data)
    est <- coef(summary(fitREML))
    ci <- confint(fitREML, level = 0.95, method = "Wald")
  out<- noquote(paste(colnames(data)[i],a, b, rm, rc,est[2,1], ci[4,1], ci[4,2],
                      sep =" "))
  #write.table(out, file = "Sang_MLPE_results_2Rules_separate.txt", append=T,row.names=F,col.names=F, quote = F) #uncomment to write file
}
```

## IBE with IBD

IBE combinations
Scale and center distances
```{r}
ibe.data<-alldata[,c("gdist","pop1","pop2","pair_name","eucdist","IBE.Zn")] #include IBD, top IBE

ibe.scaled<-scale(ibe.data[,5:ncol(ibe.data)],center=T,scale=T)
ibe.scaled<-cbind(ibe.data[,1:4],ibe.scaled)

data<-ibe.scaled
pop1<-data$pop1
pop2<-data$pop2
```
 
```{r}
#scale depends on which data are loaded above
fit <- MLPEnoREML(gdist ~ data$IBE.Zn + data$eucdist + (1|pop1), data=data)
anova(fit)
  b <- BIC(fit)
  a <- AICc(fit, k = length(unique(pop1))) #this make k = # pops, not # pairs
  r <- r.squaredGLMM(fit)
  rm <- r[1]
  rc <- r[2]
  fit
  a
  rm
fit <- MLPE(gdist ~ data$IBE.Zn + data$eucdist + (1|pop1), data=data)
  est <- coef(summary(fit))
est

#not written to file to keep separate
```

## Plot IBD. vs FST

```{r}
ggplot(data = plotdata, aes(x=(plotdata$eucdist/1000),y=plotdata$gdist,color = plotdata$Cluster)) +
  xlab("Euclidean Distance (km)") +
  ylab(expression(F[ST])) +
  geom_point(shape=16,size=1.4) +
  scale_color_manual(values = c("#00A0FA","#AA0A3C", "#000000"), name = "Sub-region", labels = c("West Chester", "State Line", "Between Regions")) +
  #geom_smooth(method=lm,color="black",size=.5,se=FALSE,linetype="solid") +
  theme_classic(base_size = 11) +
  theme(axis.line.x = element_line(color="black", size = .5),
        axis.line.y = element_line(color="black", size = .5),
        axis.title = element_text(size = 10),
        axis.text = element_text(color="black",size = 9),
        plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm")) 

#ggsave("IBD_vs_FST.pdf",plot=last_plot(),width=130,height=180,units="mm",dpi=600) #uncomment to write file
```


## Genetic diversity vs. land use

Read in data and prep for analysis
Created combined cropland and low + high intensity development
```{r}
div <- read.csv("DiversityandBufferData10Mar2020.csv", header = T)
div$cropalldev500<-div$crop500+div$alldev500
div$cropalldev1k<-div$crop1k+div$alldev1k
div$cropalldev250<-div$crop250+div$alldev250
```

Linear regressions for combined crop and development
```{r}
d<-(lm(Num ~ cropalldev500, data = div))
e<-(lm(Hs ~ cropalldev500, data = div))
f<-(lm(Gis ~ cropalldev500, data = div))

d1<-summary(d)$coefficients[2,4]
e1<-summary(e)$coefficients[2,4]
f1<-summary(f)$coefficients[2,4]

pval<-c(d1,e1,f1)

#Benjamini Hochberg adjustment
p.adjust(pval, "BH")
```

Check residuals for spatial autocorrelation
```{r}
library(ape)

cropalldev.dists <- as.matrix(dist(cbind(div$UTM_N, div$UTM_E)))
cropalldev.dists.inv <- 1/cropalldev.dists
diag(cropalldev.dists.inv) <- 0 

d.I <- Moran.I(d$residuals, cropalldev.dists.inv)
e.I <- Moran.I(e$residuals, cropalldev.dists.inv)
f.I <- Moran.I(f$residuals, cropalldev.dists.inv)

c(d.I$p.value, e.I$p.value,f.I$p.value) #none of these residuals are significantly autocorrelated
```


## Plot expected heterozygosity vs. all types of human developed land
```{r}
library(ggplot2)

ggplot(data = div, aes(x=div$cropalldev500,y=div$Hs)) +
  xlab("Developed + crop") +
  ylab(expression(H[e])) +
  geom_point(shape=16,color="black",size=1) +
  geom_smooth(method=lm,color="black",size=.5,se=FALSE,linetype="solid") +
  theme_classic(base_size = 11) +
  theme(axis.line.x = element_line(color="black", size = .5),
        axis.line.y = element_line(color="black", size = .5),
        axis.title = element_text(size = 10),
        axis.text = element_text(color="black",size = 9),
        plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm")) 

#ggsave("GenDiv_LandCov.pdf",plot=last_plot(),width=80,height=80,units="mm",dpi=600) #uncomment to write file
```



